Hierarchically Supervised Latent Dirichlet Allocation

### 2011-05-31 paper/introduction.tex

In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data  about a document;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.  These labels are usually generatively modeled as having been conditionally draw given an inferred document-specific topic mixture.  It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.   Consider web retail data.  Web retailers often have both a browse-able product hierarchy and free-text descriptions for all products they sell.   The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.  We hypothesis that hierarchical labels should, at least in theory, provide better signal than the simpler unstructured supervision previously considered.  Results from applying our model to medical record and web retail data suggests that this is likely to be the case.  In particular, we observed big gains in our primary goal of out-of-sample label prediction when using hierarchical supervision.

The remainder of this paper is structured as follows.  In Section~\ref{sec:related_work} we review related work, in Section~\ref{sec:model} we introduce hierarchically supervised LDA (HSLDA), and in Section~\ref{sec:experiments} we apply HSLDA to health care and web retail data, showing predictive performance and improved topic generation.

>>>>

In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.  These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.  It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.   Consider web retail data.  Web retailers often have both a browse-able product hierarchy and free-text descriptions for all products they sell.   The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.  We hypothesize that such hierarchical labels should, at least in theory, provide better supervision than the simpler unstructured labels previously considered.  Results from applying our model to both medical record and web retail data suggests that this is likely the case.  In particular, we observe gains in our primary goal of out-of-sample label prediction that result specifically from leveraging hierarchical supervision.

The remainder of this paper is structured as follows.   Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.
